HO unification from object language to meta language

### 2024-03-13 paper/main.tex
manual rebase of main.tex from main

The equational theory of Coq's Dependent Type Theory is very rich. In
addition to the usual $\eta\beta$-equivalence for functions, terms (hence types)
are compared up to definition unfolding and fixpoint unrolling. Still,
for efficiency and predictability reasons, most form of automatic proof search
employ a unification procedure that captures a simpler one,
just $\eta\beta$, and that solves higher-order problems
restricted to the pattern fragment $\mathcal{L}_\lambda$~\cite{miller92jsc}.
We call this unification procedure \Uo{}.

The equational theory of the meta language Elpi that we want to use to
implement a form of proof automation is strikingly similar, since it
it comprises $\eta\beta$ (for the meta language functions), and the
unification procedure \Ue{} solves higher-order problems in
$\mathcal{L}_\lambda$.

@@

\newcommand{\specunif}[3]{
  #3_i \in \mathcal{L}_\lambda \Rightarrow %
    \exists \rho, %
      \rho #3_1 #1 \rho #3_2  %
        \Leftrightarrow #3_1 #2 #3_2 \mapsto \rho' \subseteq \rho 
}

@@

Given 
$$
\C{s_1}{t_1}{l_1} \land \C{s_2}{t_2}{l_2}
$$
we define
$$
s_1 \Uo s_2 \mapsto \rho \stackrel{def}{=}
    t_1 \Ue t_2 \mapsto \sigma \land
    \D{\sigma}{l_1 + l_2}{\rho}
$$
Where $l_1 + l_2$ is the list concatenation of links.

We write $s \in \mathcal{L}_\lambda$ if all unif variables in
$s$ are applied to distinct bound variables.
%
\begin{gather}
\unifcorrect{\Eo}{\Uo}{s}\label{prop:correct}\\
\unifcomplete{\Eo}{\Uo}{s}\label{prop:complete}\\
% \forall \rho'\rho, 
  \rho s_1 \Eo \rho s_2 \Rightarrow
  \rho' \subseteq \rho \Rightarrow
  \rho's_i \in \mathcal{L}_\lambda \Rightarrow
  \rho' s_1 \Uo \rho' s_2 \label{prop:simulation}
\end{gather}

@@

Property \ref{prop:complete} is also relevant to use \Uo for logic programming.
In particular we want failures to occur as early as possible, so want
the decomp phase to take place immediately after \Ue, and fail if need be.
This becomes particularly important since compile may introduce two
ho variables for the same fo one, leaving the task of unifying the solutions
to decomp.

@@

In \Ho the representation of \coqIn{P x} is instead \elpiIn{uva N [x]}.
We say that the unification variable \elpiIn{uva N L} is in \llambda iff
\elpiIn{distinct-names L} holds.

\begin{elpicode}
type distinct-names list tm -> o.
distinct-names [].
distinct-names [X|XS] :- name X, not(mem X XS),
  distinct-names XS.
\end{elpicode}

@@

Together with the description of the terms of the language, we need some
auxiliary functions to perform operations like term equality, unification,
dereferencing. There predicates are supposed to be implemented in the OL and the
ML following respectively their specification. In the following few paragraphs
we give a brief definition of them.

@@

Since in our encoding we explicitly carry a substitution we need to
define the operation that applies it to a term. Its definition
takes care to normalize (flatten) applications, for example
it turns
\elpiIn{(app [app [con"f",con"a"],con"b"])}
into
\elpiIn{(app [con"f",con"a",con"b"])}.

@@

\begin{elpicode}
type fderef fsubst -> fm -> fm -> o.                 ~($\customlabel{pred:fderef}{\rho s}$)~
% flattening
fderef S (fapp [fuva N|A]) R :- set? N S (fapp B), !,
  append A B C, fderef S (fapp C) R.
% traversal
fderef S (fapp A) (fapp B) :- map (fderef S) A B.
fderef S (flam F) (flam G) :- 
  pi x\ fderef S x x => fderef S (F x) (G x).
fderef _ (fcon C) (fcon C).
% dereferencing
fderef S (fuva N) R :- set? N S T, fderef S T R.
fderef S (fuva N) (fuva N) :- unset? N S.
\end{elpicode}

\noindent
We use the cut operator to keep the code compact. It is possible to rewrite
the rule for application traversal so that it is mutually exclusive with the
first one, but that requires a rather verbose analysis of the head of \elpiIn{A}.

@@

\begin{elpicode}
type deref subst -> tm -> tm -> o.                   ~($\customlabel{pred:deref}{\sigma t}$)~
% ... similar to above ...
deref S (uva N A) R :- set? N S F, move F A T, deref S T R.
deref S (uva N A) (uva N B) :- unset? N S, map (deref S) A B.
type move assignment -> list tm -> tm -> o.
move (abs F) [X | L] R :- move (F X) L R.
move (val A)         [] A :- !.
move (val (app A))   L (app X) :- append A L X.
move (val (uva N A)) L (uva N X) :- append A L X.
move (val (con H))   L (app [con H | L]).
\end{elpicode}

\noindent
Note that when the substitution \elpiIn{S} maps a unification variable
\elpiIn{N} to an assignment \elpiIn{F} we ....

@@

\section[Compilation: fo\_tm to tm]{Compilation}
\label{sec:compilation}

@@

\begin{figure*}
\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Ue)} \PYG{k+kt}{tm -> tm -> subst -> subst -> o}~.
% Congruence
(app A ~\Ue~app B) S S1 :- fold2 ~(\Ue)~A B S S1.
(lam F ~\Ue~lam G) S S1 :- pi x\ (pi S\ (x ~\Ue~x) S S) => (F x ~\Ue~G x) S S1.
~\Ue~(con C) (con C) S S.
% deref
(uva N A ~\Ue~T) S S1 :- set? N S F, move F A T1, (T1 ~\Ue~T) S S1.
(T ~\Ue~uva N A) S S1 :- set? N S F, move F A T1, (T ~\Ue~T1) S S1.
% flex-flex
(uva N A ~\Ue~uva M B) S S3 :-  unset? M, unset? N,
  distinct-names A, distinct-names B, 
  new S W S1, prune W Args1 B Ass, 
  assign N S1 Ass S2, assign M S2 Ass S3.
% assignment
(uva N A ~\Ue~T) S S1 :- distinct-names A, not (T = uva _ _), not_occ N S T,  
  bind A T T1, assign N S T1 S1.
(T ~\Ue~uva N A) S S1 :- distinct-names A, not (T = uva _ _), not_occ N S T, 
  bind A T T1, assign N S T1 S1.

type distinct-names list tm -> o.
distinct-names [].
distinct-names [X|XS] :- name X, not(mem X XS),
distinct-names XS.

@@

type new memory A -> nat -> memory A -> o.
type prune .
type move .
type beta.
type not_occ.
TODO

\end{elpicode}
\caption{Implementation of the \Ue predicate for \Ho}
\label{code:ue-pred}
\Description[code:unif-pred]{code:unif-pred}
\end{figure*}

>>>>

The equational theory of Coq's Dependent Type Theory is very rich. In
addition to the usual $\eta\beta$-equivalence for functions, terms (hence types)
are compared up to proposition unfolding and fixpoint unrolling. Still,
for efficiency and predictability reasons, most form of automatic proof search
employ a unification procedure that captures a simpler one,
just $\eta\beta$, and that solves higher-order problems
restricted to the pattern fragment $\llambda$~\cite{miller92jsc}.
We call this unification procedure \Uo{}.

The equational theory of the meta language Elpi that we want to use to
implement a form of proof automation is strikingly similar, since it
it comprises $\eta\beta$ (for the meta language functions), and the
unification procedure \Ue{} solves higher-order problems in
$\llambda$.

@@

\newcommand{\specunif}[3]{
  #3_i \in \llambda \Rightarrow %
    \exists \rho, %
      \rho #3_1 #1 \rho #3_2  %
        \Leftrightarrow #3_1 #2 #3_2 \mapsto \rho' \subseteq \rho 
}

@@

We represent a logic program run for a goal g as
a list steps $p$ of length $\mathcal{N}$ each made of a unification problem
between terms $\mathcal{S}_{p_l}$ and $\mathcal{S}_{p_r}$
taken from the set of all terms $\mathcal{S}$.
If a rule is used
twice in a run we make two copies of the terms (the predicates)
occurring in the head or premises. So a run is the conjunction of all
unification problems, where $\rho_0$ is the empty substitution
%
$$
\bigwedge_{p = 1 \ldots \mathcal{N}}
\underbrace{
  \rho_{(p-1)}\mathcal{S}_{p_l} \Uo \rho_{(p-1)}\mathcal{S}_{p_l} \mapsto \rho_p
}_\mathrm{step}
$$

We simulate this run as where $\sigma_0$ is the empty substitution
%
\newcommand{\solve}{\ensuremath{\mathrm{check}}}
\begin{gather*}
\mathcal{T} \times \mathbb{L} = \{ (t_j,l_j) | s_j \in \mathcal{S}, \C{s_j}{t_j}{l_j} \}\\
\bigwedge_{p = 1 \ldots \mathcal{N}}
 \underbrace{
  \sigma_{(p-1)}\mathcal{T}_{p_l} \Ue \sigma_{(p-1)}\mathcal{T}_{p_r} \mapsto \sigma_p' \land
  \solve~(\mathbb{L},\sigma_p')\mapsto \sigma_p
 }_\mathrm{step'}\\
\D{\sigma_{\mathcal{N}}}{\mathbb{L}}{\rho_{\mathcal{N}}}
\end{gather*}

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
  $\forall p, \mathrm{step}(p,\rho) \mapsto \rho' \Leftrightarrow
              \mathrm{step'}(p,\sigma) \mapsto \sigma'$
\end{proposition}

\newcommand{\maybebeta}{\ensuremath{\stackrel{?}{\beta}}\xspace}
\begin{definition}[\maybebeta]
  $t$ is in \maybebeta if it contains a subterm matching the
  pattern \elpiIn{app[uva _ _|_]}
\end{definition}

\newcommand{\maybeeta}{\ensuremath{\stackrel{?}{\eta}}\xspace}
\begin{definition}[\maybeeta]
  $t$ is in \maybeeta if it contains a subterm matching the
  pattern \elpiIn{lam x\ .. lam y\ app[con _,x, ...,uva _ [y]]}\\
   or
  \elpiIn{lam x\lam y\uva _ [y,x]}
\end{definition}

\begin{definition}[Normal form]
  We write $norm(\mathcal{T})$ when $\mathcal{T} \cap (\maybebeta \cup \maybeeta) = \emptyset$
\end{definition}

We write $\sigma\mathcal{T} = \{ \sigma t_i | t_i \in \mathcal{T}\}$.

\begin{proposition}[Normal form preservation]
  $norm(\sigma\mathcal{T}) \land \mathrm{step'}(p,\sigma) \mapsto \sigma'
  \Rightarrow norm(\sigma\mathcal{T})$   
\end{proposition}

If we focus on a single unification problem, that is
$\mathcal{S} = \{ s_1, s_2 \}$ and 
$\C{s_1}{t_1}{l_1}$ and $\C{s_2}{t_2}{l_2}$ we write
$$
\begin{array}{l}
s_1 \Uo s_2 \mapsto \rho \stackrel{def}{=}\\
\quad    t_1 \Ue t_2 \mapsto \sigma' \land
    \solve~(\{l_1,l_2\},\sigma') \mapsto \sigma'' \land
    \D{\sigma''}{\{l_1,l_2\}}{\rho}
\end{array}
$$
\noindent
We write $s \in \llambda$ if all unif variables in
$s$ are applied to distinct bound variables.

\begin{proposition}[Properties of \Uo{}]
\begin{gather}
  \unifcorrect{\Eo}{\Uo}{s}\label{prop:correct}\\
\unifcomplete{\Eo}{\Uo}{s}\label{prop:complete}\\
% \forall \rho'\rho, 
  \rho s_1 \Eo \rho s_2 \Rightarrow
  \rho' \subseteq \rho \Rightarrow
  \rho's_i \in \llambda \Rightarrow
  \rho' s_1 \Uo \rho' s_2 \label{prop:simulation}
\end{gather}
\end{proposition}

@@

Property \ref{prop:fidelity} is also relevant to use \Uo for logic programming.
In particular we want failures to occur as early as possible, so want
the check phase to take place immediately after \Ue, and fail if need be.

\paragraph{The intuition} is that a term s is compiled in a term t where any
``problematic'' sp is replaced by a hole hp and an accessory
(suspended) unification problem tp \Ue hp (called link) where
tp is the compiled of sp. 
On t \Ue is well behaved (captures \Eo), and check solves links when it can by
using \Ue or keeps them suspended until the decompilation phase.
check also detects occur checks trough the suspended unif problems
in order to grant \cref{prop:fidelity}.

@@

In \Ho the representation of \coqIn{P x} is instead \elpiIn{uva N [x]}.
We say that the unification variable \elpiIn{uva N L} is in \llambda iff
\elpiIn{distinct L} holds.

\begin{elpicode}
type distinct list A -> o.
distinct [].
distinct [X|XS] :- name X, not(mem X XS),
  distinct XS.
\end{elpicode}

@@

Together with the description of the terms of the language, we need some
auxiliary functions to perform operations like term equality, unification,
dereferencing. There predicates are supposed to be implemented in the OL and the
ML following respectively their specification. In the following few paragraphs
we give a brief proposition of them.

@@

Since in our encoding we explicitly carry a substitution we need to
define the operation that applies it to a term. Its proposition
takes care to normalize (flatten) applications, for example
it turns
\elpiIn{(app [app [con"f",con"a"],con"b"])}
into
\elpiIn{(app [con"f",con"a",con"b"])}.

@@

\begin{elpicode}
type fder fsubst -> fm -> fm -> o.
fder S (fapp A) (fapp B) :- map (fder S) A B.
fder S (flam F) (flam G) :- 
  pi x\ fder S x x => fder S (F x) (G x).
fder _ (fcon C) (fcon C).
fder S (fuva N) R :- set? N S T, fder S T R.
fder S (fuva N) (fuva N) :- unset? N S.

type fderef fsubst -> fm -> fm -> o.                 ~($\customlabel{pred:fder}{\rho s}$)~
fderef S T R :- fder S T T', napp T' R.

type napp fm -> fm -> o.
\end{elpicode}

\noindent
TODO explain napp.
We use the cut operator to keep the code compact. It is possible to rewrite
the rule for application traversal so that it is mutually exclusive with the
first one, but that requires a rather verbose analysis of the head of \elpiIn{A}.

@@

\begin{elpicode}
type deref subst -> tm -> tm -> o.                   ~($\customlabel{pred:deref}{\sigma t}$)~
deref S (app A) (app B) :- map (deref S) A B.
deref S (lam F) (lam G) :- 
  pi x\ deref S x x => deref S (F x) (G x).
deref _ (con C) (con C).
deref S (uva N A) R :- set? N S F, move F A T, deref S T R.
deref S (uva N A) (uva N B) :- unset? N S, map (deref S) A B.
type move assignment -> list tm -> tm -> o.
move (abs F) [X | L] R :- move (F X) L R.
move (val A)         [] A :- !.
move (val (app A))   L (app X) :- append A L X.
move (val (uva N A)) L (uva N X) :- append A L X.
move (val (con H))   L (app [con H | L]).
\end{elpicode}

\noindent
TODO: no need to napp, see the beta section.
Note that when the substitution \elpiIn{S} maps a unification variable
\elpiIn{N} to an assignment \elpiIn{F} we ....

@@

\section[Compilation: fo\_tm to tm]{Basic compilation \Fo{} to \Ho{}}
\label{sec:compilation}

In this section we describe a basic compilation scheme that we refine
later, in the following sections.

The objective of the compilation is to recognize the higher-order variables
available in \Ho{} when expressed in a first order way in \Fo{}. The compiler
also generates a list of links that are used to bring back the substitution from
\Ho{} to \Fo{} and allocates in the memory a cell for each variable.

\begin{elpicode}
kind link type.
type link nat -> nat -> nat -> subst. % link Fo Ho Arity
typeabbrev links list link.
type comp fm -> tm -> links -> links -> subst -> subst -> o.
comp (fcon X) (con X) L L S S.
comp (flam F) (lam G) K L R S :- pi x y\
  (pi A S\ comp x y L L S S) => comp (F x) (G y) K L R S.
comp (fuva M) (uva N []) K [link M N z|K] R S :- new R N S.
comp (fapp[fuva M|A]) (uva N B) K L R S :- distinct A, !,
  fold4 comp A B K K R R,
  new R N S, len A Arity,
  L = [link N M Arity | K].
comp (fapp A) (app B) K L R S :- fold4 comp A B K L R S.
\end{elpicode}

\noindent
Note that \elpiIn{link} carries the arity (number of expected arguments) of
the variable. \marginpar{say when this is needed}

\begin{elpicode}
type solve-links links -> links -> subst -> subst -> o.
solve-links L L S S.
\end{elpicode}

Then decomp

\begin{elpicode}
type decompile links -> subst -> fsubst -> o.
decompile L S O :-
  map (_\r\r = none) S O1, % allocate empty fsubst
  (pi N X\ knil N X :- mem L (link X N _) ; N = X) =>
    decompl S L O1 O.
type knil nat -> nat -> o.

type decompl links -> subst -> fsubst -> o.
decompl S [] [].
decompl S [link _ N _|L] O P :- unset? N S X,
  decompl S L O P.
decompl S [link M N _|L] O P :- set? N S X,
  decomp-assignment S X T, assign M O (some T) O1,
  decompl S L O1 P.

type decomp-assignment subst -> assignment -> fm -> o.
decomp-assignment S (abs F) (flam G) :-
  pi x y\ decomp-tm S x y => decomp-assignment S (F x) (G y).
decomp-assignment S (val T) T1 :- decomp S T T1.

type decomp subst -> tm -> fm.
decomp _ (con C) (fcon C).
decomp S (app A) (app B) :- map (decomp S) A B.
decomp S (lam F) (flam G) :-
  pi x y\ decomp S x y => decomp S (F x) (G y).
decomp S (uva N A) R :- set? N S F,
  move F A T, decomp S T R.
decomp S (uva N A) R :- unset? N S,
  map (decomp S) A B, knil N M, napp (fapp[fuva M|B]) R.
\end{elpicode}

\marginpar{TODO link}
\marginpar{TODO nuove subst}

Now unif\marginpar{TODO: code unif}

\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Uo)} \PYG{k+kt}{fm -> fm -> subst -> subst -> o}~.
(X ~\Uo~Y) S S1 :-
  fderef S X X0, fderef S Y Y0,                    ~($\customlabel{step:norm}{norm}$)~
  comp X0 X1 [] S0 [] L0,                        ~($\customlabel{step:compile}{compile}$)~
  comp Y0 Y1 S0 S1 L0 L1,
  (X1 ~\Ue~Y1) [] HS0,                              ~($\customlabel{step:unify}{unify}$)~
  solve-links L1 L2 HS0 HS1,                        ~($\customlabel{step:link}{link}$)~
  decompile L2 HS1  S1.                        ~($\customlabel{step:decompile}{decompile}$)~
\end{elpicode}

\subsection{Prolog simulation}
Allows us to express the properties.
we take all terms involved in a search (if a rule is used twice we simply
take a copy of it), we compile all of them, and then we pick the unification
prblems among these terms and step trough them.

\begin{elpicode}
type pick list A -> (pair nat nat) -> (pair A A) -> o.
pick L (pr X Y) (pr TX TY) :- nth X L TX, nth Y L TY.

type prolog-fo list fm -> list (pair nat nat) -> subst -> o.
prolog-fo Terms Problems S :-
  map (pick Terms) Problems FoProblems,
  fold4 ~(\Uo)~ FoProblems [] S.

type step-ho (pair tm tm) -> links -> links -> subst -> subst -> o.
step-ho (pr X Y) L0 L1 S0 S2 :-
  (X1 ~\Ue~Y1) S0 S1,
  solve-links L0 L1 S1 S2.

type prolog-ho list fm -> list (pair nat nat) -> subst -> o.
prolog-ho Terms Problems S :-
  fold4 comp Terms HoTerms [] L0 [] HS0,
  map (pick HoTerms) Problems HoProblems,
  fold4 step-ho HoProblems L0 L HS0 HS,
  decompile L HS S.
\end{elpicode}

\noindent
the proprty is that if a step for Fo succeds then the Ho one does, and
if Fo fails then the Ho fails ()

\subsection{Example}

OK

\begin{elpicode}
Terms [ flam x\ fapp[fcon"g",fapp[fuva z, x]] 
      , flam x\ fapp[fcon"g", fcon"a"] ]
Problems = [ pr z (s z) ] % $\lambda x.g (F x) = \lambda x.g a$ 
lam x\ app[con"g",uva z [x]] ~\Uo~lam x\ app[con"g", con"a"]
link z z (s z)
HS = [some (abs x\con"a")]
S = [some (flam x\fcon a)]
\end{elpicode}

KO

\begin{elpicode}
  Terms [ flam x\ fapp[fcon"g",fapp[fuva z, x]] 
  , flam x\ fapp[fcon"g", fcon"a"] ]
Problems = [ pr 0 1   % $A = \lambda x.x$
           , pr 2 3 ] % $A a = a$ 
lam x\ app[con"g",uva z [x]] ~\Uo~lam x\ app[con"g", con"a"]
link z z (s z)
HS = [some (abs x\con"a")]
S = [some (flam x\fcon a)]
lam x\ app[f, app[X, x]] = Y,
  lam x\ x) = X.
\end{elpicode}

@@

\begin{figure*}
\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Eo)} \PYG{k+kt}{ftm -> ftm -> o}~.                           ~($\customlabel{pred:fequal}{\Eo}$)~
fapp A ~\Eo~fapp B :- map ~(\Eo)~ A B.
flam F ~\Eo~flam G :- pi x\ x ~\Eo~x => F x ~\Eo~G x.
fcon C ~\Eo~fcon C.
fuva N ~\Eo~fuva N.
flam F ~\Eo~T :-                                       ~($\customlabel{clause:eta1}{\eta_l}$)~
  pi x\ beta T [x] (R x), x ~\Eo~x => F x ~\Eo~R x.
T ~\Eo~flam F :-                                       ~($\customlabel{clause:eta2}{\eta_r}$)~
  pi x\ beta T [x] (R x), x ~\Eo~x => R x ~\Eo~F x.
fapp [flam X | L] ~\Eo~T :- beta (flam X) L R, R ~\Eo~T. ~($\customlabel{clause:beta1}{\beta_l}$)~
T ~\Eo~fapp [flam X | L] :- beta (flam X) L R, T ~\Eo~R. ~($\customlabel{clause:beta2}{\beta_r}$)~

type beta fm -> list fm -> fm -> o.
beta A [] A.
beta (flam F) [H | L] R :- subst F H B,
  beta B L R. % since F could be x\app[x|_] and H be lam _ 
beta (fapp A) L (fapp X) :- append A L X.
beta (fuva N) L (fapp [fuva N | L]).
beta (fcon H) L (fapp [fcon H | L]).

type subst (fm -> fm) -> fm -> fm -> o.
subst F H B :- napp (F H) B. % since (F H) may generate (app[app _|_])

type napp fm -> fm -> o.
napp (fcon C) (fcon C).
napp (flam F) (flam G) :- pi x\ napp x x => napp (F x) (G x).
napp (fapp[fapp L|M]) R :- !, append L M N, napp (fapp N) R.
napp (fapp[X]) R :- !, napp X R.
napp (fapp A) (fapp B) :- map napp A B.
napp (fuva N) (fuva N).
\end{elpicode}
\caption{Full implementation of the \Eo predicate for \Fo}
\label{code:eo-pred}
\Description[code:eo-pred]{code:eo-pred}
\end{figure*}

\begin{figure*}
\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Ue)} \PYG{k+kt}{tm -> tm -> subst -> subst -> o}~.
% Congruence
(app A ~\Ue~app B) R S :- fold2 ~(\Ue)~A B R S.
(lam F ~\Ue~lam G) R S :- pi x\ (pi S\ (x ~\Ue~x) S S) => (F x ~\Ue~G x) R S.
~\Ue~(con C) (con C) S S.
% deref
(uva N A ~\Ue~T) R S :- set? N S F, move F A T1, (T1 ~\Ue~T) R S.
(T ~\Ue~uva N A) R S :- set? N S F, move F A T1, (T ~\Ue~T1) R S.
% flex-flex
(uva N A ~\Ue~uva M B) S S3 :-  unset? M, unset? N,
  distinct A, distinct B, 
  new S W S1, prune W Args1 B Ass, 
  assign N S1 Ass S2, assign M S2 Ass S3.
% assignment
(uva N A ~\Ue~T) R S :- distinct A, not (T = uva _ _), not_occ N S T,  
  bind A T T1, assign N S T1 S1.
(T ~\Ue~uva N A) R S :- distinct A, not (T = uva _ _), not_occ N S T, 
  bind A T T1, assign N S T1 S1.

type distinct list A -> o.
distinct [].
distinct [X|XS] :- name X, not(mem X XS),
distinct XS.

@@

type new memory A -> nat -> memory A -> o.
new [] z [none].
new [X|XS] (s N) [X|YS] :- new XS N YS.

type prune .
type move .
type beta.
type bind.
type not_occ.
TODO

type fold2 (A -> A1 -> B -> B -> o) -> list A -> list A1 -> B -> B -> o.
fold2 _ [] [] A A.
fold2 F [X|XS] [Y|YS] A A1 :- F X Y A A0, fold2 F XS YS A0 A1.

\end{elpicode}
\caption{Implementation of the \Ue predicate for \Ho}
\label{code:ue-pred}
\Description[code:ue-pred]{code:ue-pred}
\end{figure*}

\begin{figure*}
\begin{elpicode}
type fold4 (A -> A1 -> B -> B -> C -> C -> o) -> list A -> list A1 -> B -> B -> C -> C -> o.
fold4 _ [] [] A A B B.
fold4 F [X|XS] [Y|YS] A A1 B B1 :- F X Y A A0 B B0, fold4 F XS YS A0 A1 B0 B1.

type len list A -> nat -> o.
len [] z.
len [_|L] (s X) :- len L X.

\end{elpicode}
\caption{Implementation of the compiler}
\label{code:ue-pred}
\Description[code:ue-pred]{code:ue-pred}
\end{figure*}
